Psychological Needs During Intergroup Contact: Three Experience Sampling Studies

### 2022-11-08 Methods-and-Results.tex
initial shortening

We begin by assessing the stability of our main analyses across the
three studies. To do so, we plot the effect sizes of each parameter of
interest in a forest plot. Forest plots are a popular form of evidence
synthesis and are commonly found within meta-analytic efforts. By
plotting a standardized effect size and confidence interval of key
parameters for each study, forest plots offer a visual method of
identifying broader patterns across studies. This affords us a chance to
spot stable as well as unreliable effects more easily. While not the
focus of our efforts here, forest plots are commonly accompanied with a
meta-analytic effect size that summarizes the effect size across
studies.

We assess forest plots of the basic contact hypothesis test
(\fgrref{fig:ContactHypothesis}), the core need fulfillment proposal
(\fgrref{fig:AllportNeedFulfillment}), as well as the robustness
analyses (\fgrref{fig:Robustness}). For the basic contact hypothesis
test we see that outgroup contact has a strong and consistent effect on
outgroup attitudes while interactions with non-outgroup members
consistently have no meaningful effect on outgroup attitudes. While this
would be expected from the general intergroup contact literature, this
is not a trivial finding. We are among the first to assess the contact
hypothesis using real-life intensive longitudinal data and these close
to event assessments are comparable but not equivalent to the
assessments found in cross-sectional data (also see below). When looking
at \fgrref{fig:AllportNeedFulfillment}, we see that the fulfillment of
core needs during intergroup contacts is consistently a meaningful
predictor of outgroup attitudes and interaction quality perceptions (see
\fgrref{fig:AllportNeedFulfillment} A and B). We also see that the
effect of core need fulfillment on outgroup attitudes is strongly
reduced when modeled together with interaction quality perceptions,
supporting our assertion that interaction quality offers a psychological
mechanism for the direct effect.

@@

Finally, we use the qualitative data from the participants'
self-identified core needs to contextualize the results of our main
analysis. However, because our participants jointly reported on
thousands of intergroup contacts, it would not be feasible to analyse
these qualitative responses in a traditional qualitative content
analysis. We instead rely on recent machine learning advances within the
natural language processing domain. For our analysis we use the BERT
language model. BERT (Bidirectional Encoder Representations from
Transformers) was developed by Google in 2018 and today forms a key
elements of many natural language processing workflows. In its essence,
BERT is a framework that allows users to codify every word in relation
to every other word within a large set of documents. This task is
immensely computationally intensive and most users of the framework
don't create their own model. We use the ``all-mpnet-base-v2'' language
model for our embedding. The model is a community-based version of BERT
that uses Microsoft's ``mpnet-base'' model and is fine-tuned with 1
Billion diverse sentence pairs. The model was trained with a 768
dimensional dense vector space (a relatively high number) to represent
every word. As a result, the model is widely considered to be one of the
most accurate all-purpose language models within the open-source BERT
framework. We extracted 47 topics from the 2,983 interaction goal
free-text entries --- a relatively large number of topics. The higher
number of topics allows us to retain more of the smaller topics and
leaves a relatively low number of 308 free-text entries unclassified
(10.33\%). A full write up of the topic modeling process is available in
Online Supplemental Material C.

@@

Thus, almost all other extracted topics fall into broader or narrower
need concepts that are commonly discussed within the psychological needs
literature \citep[e.g.,][]{Orehek2018a} and offer insight into a core
aspect of the migration experience that has remained broadly
under-explored \citep[][]{Kreienkamp2022d}.

>>>>

We begin by assessing the stability of our main analyses across the
three studies. To do so, we plot the effect sizes of each parameter of
interest in a forest plot. This affords us a chance to spot stable as
well as unreliable effects more easily. While not the focus of our
efforts here, forest plots are commonly accompanied with a meta-analytic
effect size that summarizes the effect size across studies.

We assess forest plots of the basic contact hypothesis test
(\fgrref{fig:ContactHypothesis}), as well as the core need fulfillment
proposal (\fgrref{fig:AllportNeedFulfillment}). For the basic contact
hypothesis test we see that outgroup contact has a strong and consistent
effect on outgroup attitudes while interactions with non-outgroup
members consistently have no meaningful effect on outgroup attitudes.
While this would be expected from the general intergroup contact
literature, this is not a trivial finding. We are among the first to
assess the contact hypothesis using real-life intensive longitudinal
data and these close to event assessments are comparable but not
equivalent to the assessments found in cross-sectional data. When
looking at the fulfillment of core needs during intergroup contacts, we
find that the motivational mechanism is consistently a meaningful
predictor of outgroup attitudes and interaction quality perceptions (see
\fgrref{fig:AllportNeedFulfillment} A and B). We also see that the
effect of core need fulfillment on outgroup attitudes is strongly
reduced when modeled together with interaction quality perceptions,
supporting our assertion that interaction quality offers a psychological
mechanism for the direct effect.

@@

Finally, we use the qualitative data from the participants'
self-identified core needs to contextualize the results of our main
analysis. However, because our participants jointly reported on
thousands of intergroup contacts, it would not be feasible to analyse
these qualitative responses in a traditional qualitative content
analysis. We instead rely on recent machine learning advances within the
natural language processing domain. For our analysis we use the BERT
language model. BERT (Bidirectional Encoder Representations from
Transformers) was developed by Google in 2018 and today forms a key
elements of many natural language processing workflows. In its essence,
BERT is a framework that allows users to codify every word in relation
to every other word within a large set of documents. We extracted 47
topics from the 2,983 interaction goal free-text entries --- a
relatively large number of topics. The higher number of topics allows us
to retain more of the smaller topics and leaves a relatively low number
of 308 free-text entries unclassified (10.33\%). A full write up of the
topic modeling process is available in Online Supplemental Material C.

@@

Thus, almost all extracted topics fall into broader or narrower need
concepts that are commonly discussed within the psychological needs
literature \citep[e.g.,][]{Orehek2018a} and offer insight into a core
aspect of the migration experience that has remained broadly
under-explored \citep[][]{Kreienkamp2022d}.
